# Explore sample

Describe the `type` variable.

```{r}
tab <- table(train$type)
data.frame(cbind(freq = tab, prop = prop.table(tab)))
```

Plot densities between `type` and the numeric variables.
Look for patterns or clusterings.

```{r densities}
numvar <- names(train)[sapply(train, class) == "numeric"]
melt(train, id.vars = "type", measure.vars = numvar) %>%
  ggplot(aes(x = type, y = value, group = type, color = type, fill = type)) +
    geom_violin(alpha = 1/2) +
    facet_wrap(~ variable) +
    theme_bw() +
    theme(legend.position = "none")
```

Show summary statistics for numeric variables.

```{r}
select(train, matches(paste(numvar, collapse = "|"))) %>% summary()
```

Show the standard deviations of the numeric variables.

```{r}
select(train, matches(paste(numvar, collapse = "|"))) %>% var() %>% diag() %>% sqrt()
```

The numeric variables are pre-scaled with values between `r paste(select(train, matches(paste(numvar, collapse = "|"))) %>% range(), collapse = ", ") %>% sprintf("(%s)")`.
Therefore, no preprocessing is needed.

Plot bivariate densities between numeric variables.
Look for correlations to reduce dimensionality of data.

```{r heatmapCorr}
corr <-
  select(train, matches(paste(numvar, collapse = "|"))) %>%
  cor()
contHighCorr <- colnames(corr)[findCorrelation(corr)]
sprintf("Remove variable due to high pair-wise correlation with other variables: %s",
        contHighCorr)
replace(corr, which(upper.tri(corr, diag = TRUE)), NA) %>%
  melt(na.rm = TRUE) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value)) + 
    geom_tile(color = "white") +
    scale_fill_gradient2(low = "blue",
                         high = "red",
                         mid = "white",
                         midpoint = 0,
                         limit = c(-1, 1),
                         space = "Lab",
                         name="R") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          panel.grid.major = element_blank()) +
    coord_fixed()
round(corr, 2)
```

Examine association between `color` and `type`.

```{r}
table(train$color)
table(train$color, train$type) %>%
  prop.table(margin = 1)
```
